Bootstrap Confidence Intervals for Sharp Regression Discontinuity Designs with the Uniform Kernel

### 2016-04-27 main_paper.tex
Fix minor notational error in theoretical section
Make minor typographical and notational tweaks to monte carlo and empirical sections

Since the conditions for identification only need to hold in a neighborhood
around the cutoff, $\mu_+$ and $\mu_-$ can be estimated by extrapolating from a
local polynomial regression.  We will focus here on local linear
regression.\footnote{%
  See \cite{HTV2001}, \cite{Porter03} or \cite{FanGijbels92} for discussions of
  the properties of local polynomial regressions for boundary problems. The
  bootstrap algorithm proposed in this paper can be extended to accommodate
  higher order polynomials discontinuities in the derivatives of the conditional
  expectation, like ``Kink RD'' design \citep{card2009b}.} %
For this model, if $h$ represents a bandwidth parameter, the estimator of
$\tau$, $\hat\tau(h)$, is defined as
\begin{equation*}
  \hat{\tau}(h) = \hat {\mu}_{+}(h) -\hat{\mu}_{-}(h)
\end{equation*}
with
\begin{align*}
  \hat {\mu}_{+}(h)
  &= \argmin_{\beta_0} \min_{\beta_1} \sum_{i=1}^{n}
  \1\{h > X_{i} \geq 0\} (Y_{i} - \beta_0 - X_{i} \beta_1)^{2}
\intertext{and}
  \hat {\mu}_{-}(h)
  &= \argmin_{\beta_0} \min_{\beta_1} \sum_{i=1}^{n}
  \1\{0 > X_{i} > -h \} (Y_{i} - \mu - X_{i} \beta)^{2}.
\end{align*}
Conventional (naive) confidence intervals can be calculated by using an
asymptotic approximation for $\hat\tau(h)$. In particular, if
\begin{gather}
  \label{eq:1}
  \frac{\hat{\tau}(h)-\tau}{\sqrt{V(h)}} \to^d N(0,1),
  \intertext{with}
  \notag
  V(h) \big/ \V(\hat\tau(h) \mid X_{1},\dots,X_{n}) \to^p 1
\end{gather}
then valid confidence intervals can be constructed through the usual method of
inverting the $t$-test.\footnote{%
  The mathematical appendix of this paper gives a precise definition for
  $V(h)$.} %
This procedure gives the widely-used interval estimator
\begin{equation*}
  \hat{\tau}(h) \pm q_{1-\alpha/2} V(h)^{1/2}
\end{equation*}
where $q_{1 - \alpha/2}$ is the $1 - \alpha/2$ quantile of the standard normal
distribution.

@@

This section presents evidence from Monte Carlo simulations that the bootstrap
procedures proposed in Section~\ref{boot} produce valid, robust confidence
intervals similar to the ones obtained by the analytical procedures established
in CCT. The bootstrap CIs obtained compare favorably to the analytical
alternative, with coverage slightly closer to nominal coverage and shorter
length of the intervals in the specifications implemented.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{lrr@{, }rr@{}rlr}
    \toprule
                      & ATE     & \multicolumn{2}{r}{95\% CI}           & $h$&      & $b$   & p-value \\
    \midrule
    LM (2007)         & --1.895 & (--3.930                    & 0.139)   &  9&     &       & 0.036   \\
    LM (2007)         & --1.198 & (--2.561                    & 0.165)   & 18&     &       & 0.081   \\
    LM (2007)         & --1.114 & (--2.138                    & --0.090) & 36&     &       & 0.027   \\
    CCT               & --3.795 & (--7.037                    & --0.554) &  3&.888 & 6.807 &         \\
    Resid.\ bootstrap & --3.792 & (--6.512                    & --0.262) &  3&.888 & 6.807 &         \\
    \bottomrule
  \end{tabular}
  \caption{The effect of Head Start assistance on mortality. The first three
    rows come from Table 3 in \cite{ludwig2007} except ``95\% CI,'' which is
    calculated using the conventional asymptotic interval estimator.
    The last two rows list results from two robust procedures.}
  \label{tbl:2}
\end{table}

@@

Table \ref{tbl:2} shows the results of the Head Start program on mortality of
children five to nine years of age.\footnote{%
  We, like Ludwig and Miller, focus on the 1973--1983 period.} %
Instead of choosing an optimal bandwidth, \cite{ludwig2007} adopted three
candidate bandwidths 9, 18, and 36. Their estimates indicate that Head Start
assistance lowers the targeted mortality rate by --1.895, --1.198 and --1.114
respectively, which are not very sensitive to the choice of bandwidths in this
range. These ATEs are also significantly different from zero (with p-value
0.036, 0.081 and 0.027) based on Ludwig and Miller's percentile-$T$ bootstrapped
p-value. The statistical inference changes when conventional analytical
confidence interval is used, which includes zero for bandwidth 9 and 18.

Results from the two robust procedures are similar to each other but greatly
differ from the original estimates. The estimated ATEs are as high as -3.795
from CCT and -3.792 from the bootstrap. Both are significantly different from
zero, though the confidence intervals are also very wide, which is likely to be
due to the much smaller bandwidths used relative to the other estimators
($h=3.888, b=6.807$).

Table \ref{tbl:3} presents the effect of the program on schooling for the cohort
aged 18-24 in 1990. The measurement of schooling is the fraction of people with
high school or more in Panel A and the fraction of people with some college or
more in Panel B. A bandwidth of 7 is used in \cite{ludwig2007}. Their estimates
suggest that Head Start assistance increases the fraction of people with high
school or more by 3\% and the fraction of people with some college or more by
3.7\%. Both are significantly different from zero based on both their p-value
and the conventional analytical confidence interval.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{lrr@{, }rr@{}rlr}
    \toprule
              & ATE   & \multicolumn{2}{r}{95\% CI} & $h$    &   & $b$  & p-value     \\
    \midrule                                                                          \\
    \multicolumn{5}{l}{Fraction ``high school or more'' (Panel A)}                    \\
    \midrule
    LM (2007) & 0.030 & (0.003                      & 0.057) & 7&     &  & 0.032    \\
    CCT       & 0.055 & (0.014                      & 0.096) & 3&.671 & 8.618       \\
    Bootstrap & 0.054 & (0.013                      & 0.096) & 3&.671 & 8.618       \\\\
    \multicolumn{5}{l}{Fraction ``some college or more'' (Panel B)}                   \\
    \midrule
    LM (2007) & 0.037 & (0.002                      & 0.073) & 7&     &  & 0.032 \T \\
    CCT       & 0.051 & (0.004                      & 0.099) & 5&.076 & 10.251      \\
    Bootstrap & 0.052 & (0.001                      & 0.094) & 5&.076 & 10.251      \\
    \bottomrule
  \end{tabular}
  \caption{The effect of Head Start assistance on education for cohort
    18-24 in 1990. Panel A uses the fraction of people with high school
    or more as dependent variable. Panel B uses the fraction of people
    with some college or more as dependent variable. The first row in
    each panel comes from Table 4 in \cite{ludwig2007} except ``95\% CI,''
    which is calculated using the conventional asymptotic interval estimator.
    The last two rows in
    each panel list results from two robust procedures.}
  \label{tbl:3}
\end{table}

>>>>

Since the conditions for identification only need to hold in a neighborhood
around the cutoff, $\mu_+$ and $\mu_-$ can be estimated by extrapolating from a
local polynomial regression.  We will focus here on local linear
regression.\footnote{%
  See \cite{HTV2001}, \cite{Porter03} or \cite{FanGijbels92} for discussions of
  the properties of local polynomial regressions for boundary problems. The
  bootstrap algorithm proposed in this paper can be extended to accommodate
  higher order polynomials discontinuities in the derivatives of the conditional
  expectation, like ``Kink RD'' design \citep{card2009b}.} %
For this model, if $h$ represents a bandwidth parameter, the estimator of
$\tau$, $\hat\tau(h)$, is defined as
\begin{equation*}
  \hat{\tau}(h) = \hat {\mu}_{+}(h) -\hat{\mu}_{-}(h)
\end{equation*}
with
\begin{align*}
  \hat {\mu}_{+}(h)
  &= \argmin_{\beta_0} \min_{\beta_1} \sum_{i=1}^{n}
  \1\{h > X_{i} \geq 0\} (Y_{i} - \beta_0 - X_{i} \beta_1)^{2}
\intertext{and}
  \hat {\mu}_{-}(h)
  &= \argmin_{\beta_0} \min_{\beta_1} \sum_{i=1}^{n}
  \1\{0 > X_{i} > -h \} (Y_{i} - \beta_0 - X_{i} \beta_1)^{2}.
\end{align*}
Conventional (naive) confidence intervals can be calculated by using an
asymptotic approximation for $\hat\tau(h)$. In particular, if
\begin{gather}
  \label{eq:1}
  \frac{\hat{\tau}(h)-\tau}{\sqrt{V(h)}} \to^d N(0,1),
  \intertext{with}
  \notag
  V(h) \big/ \V(\hat\tau(h) \mid X_{1},\dots,X_{n}) \to^p 1
\end{gather}
then valid confidence intervals can be constructed through the usual method of
inverting the $t$-test.\footnote{%
  The mathematical appendix of this paper gives a precise definition for
  $V(h)$.} %
This procedure gives the widely-used interval estimator
\begin{equation*}
  \hat{\tau}(h) \pm q_{1-\alpha/2} V(h)^{1/2}
\end{equation*}
where $q_{1 - \alpha/2}$ is the $1 - \alpha/2$ quantile of the standard normal
distribution.

@@

This section presents evidence from Monte Carlo simulations that the bootstrap
procedures proposed in Section~\ref{boot} produce valid, robust confidence
intervals similar to those obtained by the analytical procedures established
in CCT. The bootstrap CIs obtained compare favorably to the analytical
alternative, with coverage slightly closer to nominal coverage and shorter
length of the intervals in the specifications implemented.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{lrr@{, }rr@{}rrr}
    \toprule
                      & ATE     & \multicolumn{2}{r}{95\% CI}            &     &  $h$ &   $b$ & p-value \\
    \midrule
    LM (2007)         & --1.895 & (--3.930                    & 0.139)   & 9   &      &       & 0.036 \\
    LM (2007)         & --1.198 & (--2.561                    & 0.165)   & 18  &      &       & 0.081 \\
    LM (2007)         & --1.114 & (--2.138                    & --0.090) & 36  &      &       & 0.027 \\
    CCT               & --3.795 & (--7.037                    & --0.554) & 3   & .888 & 6.807 &       \\
    Resid.\ bootstrap & --3.792 & (--6.512                    & --0.262) & 3   & .888 & 6.807 &       \\
    \bottomrule
  \end{tabular}
  \caption{The effect of Head Start assistance on mortality. The first three
    rows come from Table 3 in \cite{ludwig2007} except ``95\% CI,'' which is
    calculated using the conventional asymptotic interval estimator.
    The last two rows list results from two robust procedures.}
  \label{tbl:2}
\end{table}

@@

Table \ref{tbl:2} shows the results of the Head Start program on mortality of
children five to nine years of age.\footnote{%
  We, like Ludwig and Miller, focus on the 1973--1983 period.} %
Instead of choosing an optimal bandwidth, \cite{ludwig2007} adopted three
candidate bandwidths 9, 18, and 36. Their estimates indicate that Head Start
assistance lowers the targeted mortality rate by --1.895, --1.198 and --1.114
respectively, which are not very sensitive to the choice of bandwidths in this
range. These ATEs are also significantly different from zero (with p-value
0.036, 0.081 and 0.027) based on Ludwig and Miller's percentile-$t$ bootstrapped
p-value. The statistical inference changes when conventional analytical
confidence interval is used, which includes zero for bandwidth 9 and 18.

Results from the two robust procedures are similar to each other but greatly
differ from the original estimates. The estimated ATEs are as high as --3.795
from CCT and --3.792 from the bootstrap. Both are significantly different from
zero, though the confidence intervals are also very wide, which is likely to be
due to the much smaller bandwidths used relative to the other estimators
($h=3.888$, $b=6.807$).

Table \ref{tbl:3} presents the effect of the program on schooling for the cohort
aged 18--24 in 1990. The measurement of schooling is the fraction of people with
high school or more in Panel A and the fraction of people with some college or
more in Panel B. A bandwidth of 7 is used in \cite{ludwig2007}. Their estimates
suggest that Head Start assistance increases the fraction of people with high
school or more by 3\% and the fraction of people with some college or more by
3.7\%. Both are significantly different from zero based on both their p-value
and the conventional analytical confidence interval.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{lrr@{, }rr@{}rrr}
    \toprule
              & ATE   & \multicolumn{2}{r}{95\% CI} & & $h$     & $b$ & p-value     \\
    \midrule                                                                          \\
    \multicolumn{5}{l}{Fraction ``high school or more'' (Panel A)}                    \\
    \midrule
    LM (2007) & 0.030 & (0.003                      & 0.057) & 7&     &  & 0.032    \\
    CCT       & 0.055 & (0.014                      & 0.096) & 3&.671 & 8.618       \\
    Bootstrap & 0.054 & (0.013                      & 0.096) & 3&.671 & 8.618       \\\\
    \multicolumn{5}{l}{Fraction ``some college or more'' (Panel B)}                   \\
    \midrule
    LM (2007) & 0.037 & (0.002                      & 0.073) & 7&     &  & 0.032 \T \\
    CCT       & 0.051 & (0.004                      & 0.099) & 5&.076 & 10.251      \\
    Bootstrap & 0.052 & (0.001                      & 0.094) & 5&.076 & 10.251      \\
    \bottomrule
  \end{tabular}
  \caption{The effect of Head Start assistance on education for cohort
    18--24 in 1990. Panel A uses the fraction of people with high school
    or more as dependent variable. Panel B uses the fraction of people
    with some college or more as dependent variable. The first row in
    each panel comes from Table 4 in \cite{ludwig2007} except ``95\% CI,''
    which is calculated using the conventional asymptotic interval estimator.
    The last two rows in
    each panel list results from two robust procedures.}
  \label{tbl:3}
\end{table}
